Quantum Hoare Type Theory

### 2020-06-09 ms-paper/ms-paper.tex
Switch to Pierce & Turner (2000) notation for typing
The use of Rightarrow, etc is justified in binding because we want
to treat assignment differently from equality, like in Haskell. See
http://www.haskellforall.com/2017/07/demistifying-haskell-assignment.html

We use Pierce and Turner to avoid confusion with monading binding
notation even though, <= and => are way more common in papers since
2008, see https://arxiv.org/pdf/1908.05839v1.pdf#page=32 and sec. 11
in the same survey paper by Dunfield and Krishnaswami (2019).

Further, as pointed out in the same section, "vertical arrows are
uncle

\newcommand{\chkcolor}{dBlue}
\newcommand{\syncolor}{dRed}
\newcommand{\chk}{\,\textcolor{\chkcolor}{\Leftarrow}\,}
\newcommand{\uncoloredsyn}{{\Rightarrow}}
\newcommand{\syn}{\,\textcolor{\syncolor}{\uncoloredsyn}\,}

@@

Note that the blue colored $\chk$ symbol we use in the typing rules lives in a different syntactic category from the $\Leftarrow$ symbol used in the syntax (\cref{fig:syntax}) that binds the output of primitive commands.

@@

\begin{minipage}{0.95\linewidth}
\begin{lstlisting}[language=QHaskell]
hqw : {emp} r : Bool {emp /\ Id(r, false)}
    = do q <= mkQbit false;
         measQbit q
\end{lstlisting}
\end{minipage}

@@

\begin{minipage}{0.95\linewidth}
\begin{lstlisting}[language=QHaskell]
hqw : {emp} r : Bool {emp /\ Id(r, false)}
-- P0: emp
    = do q <= mkQbit false;
-- P1: P0 \o (q |-> |0\>)
         measQbit q
-- P2: P1 \o ((q |-> -) -o emp)
\end{lstlisting}
\end{minipage}

@@

\begin{minipage}{0.95\linewidth}
\begin{lstlisting}[language=QHaskell]
rnd : {emp} r : Bool {emp}
    = do q <= mkQbit false;
         applyU (H q);
         measQbit q
\end{lstlisting}
\end{minipage}

@@

\begin{minipage}{0.95\linewidth}
\begin{lstlisting}[language=QHaskell]
rnd : {emp} r : Bool {emp}
-- P0: emp
    = do q <= mkQbit false;
-- P1: P0 \o (q |-> |0\>)
         applyU (H q);
-- P2: P1 \o ((q |-> |0\>) -o (q |-> |+\>))
         measQbit q
-- P3: P2 \o ((q |-> -) -o emp)
\end{lstlisting}
\end{minipage}

@@

\begin{minipage}{0.95\linewidth}
\begin{lstlisting}[language=QHaskell]
testBell : {emp} (a, b) : (Bool, Bool) {emp /\ Id(a, b)}  -- a,b \in {true, false}
         = do qa <= mkQbit false;
              applyU (H qa);
              qb <= mkQbit false;
              applyU (ifQ qa (X qb));
              (measQbit qa, measQbit qb)
\end{lstlisting}
\end{minipage}

@@

\begin{lstlisting}[language=QHaskell]
testBell : {emp} (a, b) : (Bool, Bool) {emp /\ Id(a, b)}
              -- P0: emp
         = do qa <= mkQbit false;
              -- P1: P0 \o (qa |-> |0\>)
              applyU (H qa);
              -- P2: P1 \o ((qa |-> |0\>) -o (qa |-> |+\>))
              qb <= mkQbit false;
              -- P3: P2 \o (qb |-> |0\>)
              applyU (ifQ qa (X qb));
              -- P4: P3 \o ((qa |-> |+\>, qb |-> |0\>) -o (qa, qb) |-> |\Phi+\>)
              (measQbit qa, measQbit qb)
              -- P5: P4 \o ((qa |-> -) -o emp) \o ((qb |-> -) -o emp)
\end{lstlisting}

@@

\subsection{Testing modular Bell pair}
\begin{lstlisting}[language=QHaskell]
testBell : {emp} (a, b) : (Bool, Bool) {emp /\ Id(a, b)}
         = do (qa, qb) <- bell;
              (measQbit qa, measQbit qb)
\end{lstlisting}

>>>>

\newcommand{\chkcolor}{dBlue}
\newcommand{\syncolor}{dRed}
\newcommand{\chk}{\,\textcolor{\chkcolor}{{\stackrel{\leftarrow}{\in}}}\,}
\newcommand{\uncoloredsyn}{{{\stackrel{\rightarrow}{\in}}}}
\newcommand{\syn}{\,\textcolor{\syncolor}{\uncoloredsyn}\,}

@@

@@

\begin{minipage}{0.95\linewidth}
\begin{lstlisting}[language=QHaskell]
hqw : {emp} r : Bool {emp /\ Id(r, false)}
    = do q <= mkQbit false;
         return measQbit q
\end{lstlisting}
\end{minipage}

@@

\begin{minipage}{0.95\linewidth}
\begin{lstlisting}[language=QHaskell]
hqw : {emp} r : Bool {emp /\ Id(r, false)}
-- P0: emp
    = do q <= mkQbit false;
-- P1: P0 \o (q |-> |0\>)
         return measQbit q
-- P2: P1 \o ((q |-> -) -o emp)
\end{lstlisting}
\end{minipage}

@@

\begin{minipage}{0.95\linewidth}
\begin{lstlisting}[language=QHaskell]
rnd : {emp} r : Bool {emp}
    = do q <= mkQbit false;
         applyU (H q);
         return measQbit q
\end{lstlisting}
\end{minipage}

@@

\begin{minipage}{0.95\linewidth}
\begin{lstlisting}[language=QHaskell]
rnd : {emp} r : Bool {emp}
-- P0: emp
    = do q <= mkQbit false;
-- P1: P0 \o (q |-> |0\>)
         applyU (H q);
-- P2: P1 \o ((q |-> |0\>) -o (q |-> |+\>))
         return measQbit q
-- P3: P2 \o ((q |-> -) -o emp)
\end{lstlisting}
\end{minipage}

@@

\begin{minipage}{0.95\linewidth}
\begin{lstlisting}[language=QHaskell]
testBell : {emp} (a, b) : (Bool, Bool) {emp /\ Id(a, b)}  -- a,b \in {true, false}
         = do qa <= mkQbit false;
              applyU (H qa);
              qb <= mkQbit false;
              applyU (ifQ qa (X qb));
              return (measQbit qa, measQbit qb)
\end{lstlisting}
\end{minipage}

@@

\begin{lstlisting}[language=QHaskell]
testBell : {emp} (a, b) : (Bool, Bool) {emp /\ Id(a, b)}
              -- P0: emp
         = do qa <= mkQbit false;
              -- P1: P0 \o (qa |-> |0\>)
              applyU (H qa);
              -- P2: P1 \o ((qa |-> |0\>) -o (qa |-> |+\>))
              qb <= mkQbit false;
              -- P3: P2 \o (qb |-> |0\>)
              applyU (ifQ qa (X qb));
              -- P4: P3 \o ((qa |-> |+\>, qb |-> |0\>) -o (qa, qb) |-> |\Phi+\>)
              return (measQbit qa, measQbit qb)
              -- P5: P4 \o ((qa |-> -) -o emp) \o ((qb |-> -) -o emp)
\end{lstlisting}

@@

\subsection{Testing modular Bell pair}
\begin{lstlisting}[language=QHaskell]
testBell : {emp} (a, b) : (Bool, Bool) {emp /\ Id(a, b)}
         = do (qa, qb) <- bell;
              return (measQbit qa, measQbit qb)
\end{lstlisting}
